---
title: "Signatures_of_Neurogensis"
author: "Brian Schilder"
date: "10/1/2017"
output: 
  html_document: 
    toc: yes
---

* **PROJECT GOAL**
  + Utilize brain-wide gene expression databases to identify regions that contain molecular signatures of adult neurogenesis.

* **DATASETS**
  1. ABA Mouse ISH
  2. ABA Macaque microarray adult/development
  3. ABA Human microarray adult/development
  4. ABA Human RNA-seq adult/development
  5. sc-RNAs-seq reference datasets for deconvolution:
    i) Nathan Skene's mouse meta-data
    ii) Neurogenic cells (Shin et al. 2015)
    iii) Human data (Darmanis et al. 2015)

____
____

* **Jeremy Miller Ideas:**
1. Using Table S1 (or some other curated list), search the ABA for "neurons in waiting."

2. Extend the analysis to include human brain.  I'm not sure if the gene list from figure 5 in the paper you attached could be used directly (since the samples we collected from human were whole dentate gyrus, if I am remembering correctly), but that would make a good starting point.

3. A similar option would be to do a deconvolution analysis on RNA-Seq data from the BrainSpan atlas using the overlapping mouse/NHP[/human?] neurogenesis list as marker genes for SGZ.  In theory we could show the relative decrease in SGZ neurons over time and also show that it doesn't reach 0 (as long as we get any reads from these genes in the oldest donors).

* **Additional Ideas:**
1. Use expression values for all genes in the SGZ (mouse or macaque) or  sc-RNA-seq from neural stem cell, and compare to transcriptome expression from each brain region in an ABA atlas.
  - Very rough. Characterizes SGZ as a whole, not necessarilly what makes it neurogenic.
  
2. Look for similar logFC of gene sets for each 'likelyCellType' (Miller et al. 2013) (e.g. immature neuron, dividing cell) in one of the above atlases (after differential expression).
  i) Can compare directly within respective species' ABA atlases
  ii) Becomes interspecies when comparing to ABA human (likely diverged)
  
3. Using mouse or macaque modules (Miller et al. 2013) (e.g. tan), identify similar modules (based on similar kME values, or gene list overlap) one of the above atlases (after WGNCA) and then find which brain structures express them the most.
  i) Problem: tan module was found by only analyzing 24 SGZ/GCL samples. Unlikely to recover the same tan module with full ABA macaque dataset.

4. Use sc-RNA-seq gene list to find similar patterns of expression across any of the ABA datasets.

5. Deconvolution:
  i) EWCE: Using a sc-RNA-seq reference dataset and an ABA atlas.
  ii) Other methods: Use a gene list as a reference  
____
____

# Import Datasets
```{r}
library(pander)
setwd("~/Desktop/Research/Signatures_of_Neurogensis/")
```

## ABA data: Mouse
```{r}
# Mouse SGZ & GCL: GSE39697
```


## ABA data: Adult Human Microarray
```{r,message=F,warning=F, eval=F}
load("/Users/schilder/Desktop/Dissertation/Gene_Expression/Raw data/Microarray/Full_Human_Microarray/All_ABA_humans.mni.RData")
# Set up eset
  sampleInfo.hum <-subset(sampleInfo.hum_all, select= -c(mni_x,mni_y,mni_z))

# Make all cols factors, & Reorder names alphabetically
  sampleInfo.hum <- lapply(sampleInfo.hum, factor)
  sampleInfo.hum <- sampleInfo.hum[order(names(sampleInfo.hum))]

library(annotate)
# Set up phenotype data
  pdata <- AnnotatedDataFrame(data.frame(sampleInfo.hum))
  sampleNames(pdata) <- sampleInfo.hum$unique_id
# Create ExpressionSet
  ABA_hum <- ExpressionSet(datExpr.hum_all, phenoData = pdata) #Correct this annot
# Add treatment condition
  ABA_hum$treatment <- "Human"
```

## ABA data: Developmental Macaque Microarray
### Macaque Preprocessing
```{r,message=F,warning=F, eval=F}
library(WGCNA)
#mac_expression <- read.csv("Full Macaque Microarray/expression_matrix.csv", header=FALSE, row.names=1)
#mac_sampleinfo <- read.csv("Full Macaque Microarray/columns_metadata.csv", header=TRUE)
#mac_probes <- read.csv("Full Macaque Microarray/rows_metadata.csv", header=TRUE)
#save(mac_expression, mac_sampleinfo, mac_probes, file="Macaque Data.RData")
  # Saved as R objects to speed up importing data
load("~/Desktop/Dissertation/Gene_Expression/Raw data/Microarray/Macaque Data.RData")

datExpr <- mac_expression
probeInfo <- mac_probes
sampleInfo <- mac_sampleinfo
rm(mac_expression,mac_probes,mac_sampleinfo)
# Select only probes with associated entrez genes
  rownames(datExpr) <- probeInfo$probe_name
  validProbes = which(probeInfo$entrez_id>=0,arr.ind=TRUE)
  datExpr <- datExpr[validProbes,]
  probeInfo <- probeInfo[validProbes,]
# Some genes have more than one probe. The collapseRows function chooses a single probe to represent a gene.
probeInfo$gene_symbol <-gsub("[()]","",probeInfo$gene_symbol) # Gets rid of weird parentheses
  probeNames = probeInfo[which(probeInfo$probe_name == rownames(datExpr)),"gene_symbol"]
  probeIDs = rownames(datExpr)
  datCollapsed = WGCNA::collapseRows(datExpr, probeNames, probeIDs) # Collapse
  datExpr <- datCollapsed$datETcollapsed
  probeInfo <- probeInfo[which(datCollapsed$selectedRow==T, arr.ind=T),]
save(datExpr, probeInfo, sampleInfo,file="ABA.macaque_preprocessed.v4.RData")
```

### Convert to eset
```{r,message=F,warning=F}
# From ABA website
load("~/Desktop/Research/Signatures_of_Neurogensis/ABA.macaque_preprocessed.v4.RData")
# From Jeremy's Dropbox
#load("~/Desktop/Research/Signatures_of_Neurogensis/neurogenesisPaperData_ComBat_5_29_12.RData")

datExpr.mac <-datExpr
probeInfo.mac <-probeInfo
sampleInfo.mac <-sampleInfo
rm(datExpr,probeInfo,sampleInfo)
library(annotate)
# Create eset
ex_mac <- ExpressionSet(datExpr.mac)
# Create macaque phenoData:
sampleInfo.mac$unique_id <- factor(paste(sampleInfo.mac$donor_name, sampleInfo.mac$structure_acronym, sampleInfo.mac$well_id, sep="_"))

p <- data.frame(sampleInfo.mac)
rownames(p) <- sampleInfo.mac$unique_id
pdata <- AnnotatedDataFrame(p)
# Replace phenoData in mac ExpressionSet
phenoData(ex_mac) <- pdata
# Insert condition
ex_mac$treatment <- "Macaque"

ex_mac <-ex_mac[,colSums(is.na(exprs(ex_mac))) == 0] # Remove cols with NA 

rm(datExpr.mac,probeInfo.mac,sampleInfo.mac,p,pdata)
```

## Import Miller et al. 2013 gene sets
* Miller J a, Nathanson J, Franjic D, Shim S, Dalley R a, Shapouri S, et al.: Conserved molecular signatures of neurogenesis in the hippocampal subgranular zone of rodents and primates. [Internet]. Development 2013;140:4633–44. 
```{r,message=F,warning=F}
library(readxl); library(annotate)
# S1: Genes differentially expressed between mouse SGZ and GCL
S1 <- data.frame(read_excel("~/Desktop/Research/Signatures_of_Neurogensis/Miller_2013/DEV097212 TableS1.xlsx",na = "NA"))
# Get rid of stars
S1$MouseGene <-gsub("[*]","",S1$MouseGene)

unique(S1$likelyCellType)
  ## Get gene list
  immature.neuron <- S1[S1$likelyCellType=="Immature Neuron",]$MouseGene
  dividing.cell <- S1[S1$likelyCellType=="Dividing Cell",]$MouseGene
  S1.genes <- S1$MouseGene
  # Convert from mouse to human genes
  library(EWCE)
  data("mouse_to_human_homologs")
  m2h = unique(mouse_to_human_homologs[,c("HGNC.symbol","MGI.symbol")])
  s1.genes <- unique(m2h[m2h$MGI.symbol %in% S1.genes,"HGNC.symbol"])
  
  paste(sum(s1.genes %in% row.names(exprs(ex_mac))),"/",length(S1.genes),
          "'Mouse SGZ-enriched' genes found in ABA Macaque microarray dataset.")

# S3: Mouse network modules
mouse_modules <- data.frame(read_excel("~/Desktop/Research/Signatures_of_Neurogensis/Miller_2013/DEV097212 TableS3.xlsx",na = "NA"))
  ## Subset specific modules
  ### tan module was associated with "neurogenesis" in Miller et al. 2013.
Mouse_mod.tan <- subset(mouse_modules, Module=="tan")$HumanOrtholog
  paste(sum(Mouse_mod.tan %in% row.names(exprs(ex_mac))),"/",length(Mouse_mod.tan),"Mouse Tan module genes found in ABA Macaque microarray dataset.")
  
# S4: Macaque network modules
macaque_modules <- data.frame(read_excel("~/Desktop/Research/Signatures_of_Neurogensis/Miller_2013/DEV097212 TableS4.xlsx", na="NA"))
  ## Subset specific modules
  ### tan module was associated with "neurogenesis" in Miller et al. 2013.
  Mac_mod.tan <- subset(macaque_modules, Module=="tan")$Gene
  paste(sum(Mac_mod.tan %in% row.names(exprs(ex_mac))),"/",length( Mac_mod.tan),"Macaque Tan module genes found in ABA Macaque microarray dataset.")
```

### Identify Macaque SGZ-enriched genes
```{r,message=F,warning=F}
# Perform DGE on Macaque SGZ vs. GCL
  SGZ.samples <-ex_mac[,pData(ex_mac)$structure_acronym==c("DGsg") ]
  GCL.samples <-ex_mac[,pData(ex_mac)$structure_acronym==c("DGgr") ]

# Limma
  library(limma); library(statmod)
  limma.data <-cbind(exprs(SGZ.samples), exprs(GCL.samples))
  fit <-lmFit(limma.data, design=
                 c(rep(1,ncol(exprs(SGZ.samples))), rep(-1,ncol(exprs(GCL.samples)))) )
  fit <- eBayes(fit)
  SGZ.vs.GCL.results <- topTable(fit, number=Inf)
  
# Pair-wise t-test
 # t.test(exprs(SGZ.samples), exprs(GCL.samples), paired=T)

sig.DEGS <-row.names(subset(SGZ.vs.GCL.results, adj.P.Val<=0.05))
paste("There are",length(sig.DEGS),"significant DEGs between Macaque SGZ and GCL.")
# Top logFC genes
topTable(fit, number=10, sort.by = "logFC")
```



# __Candidate Genes__
```{r,message=F,warning=F}
library(dplyr); library(ggplot2)
candidate.genes <- ex_mac[row.names(exprs(ex_mac)) %in% c("SOX11","SOX4","DCX","MKI67")][,pData(ex_mac)$age=="48 mo"]

copy.mac <- data.frame(t(data.frame(exprs(candidate.genes))))
#copy.mac$Region <-sub(".*_ *(.*?) *_.*", "\\1", colnames(exprs(candidate.genes)))
copy.mac$structure_name <-pData(candidate.genes)$structure_name
copy.mac$structure_acronym <-pData(candidate.genes)$structure_acronym
candidate.data <-copy.mac %>% group_by(structure_name,structure_acronym) %>% summarise(SOX11=mean(SOX11), DCX=mean(DCX), MKI67=mean(MKI67)) %>% data.frame()



# SOX11
data <-candidate.data[order(-candidate.data$SOX11),]
## All regions
ggplot(data) + geom_bar(aes(x=structure_name, y=SOX11, fill=SOX11),stat='identity') + theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(hjust=0.5))
# Top 10 expression
ggplot(rbind(subset(data,structure_acronym!="DGsg" & structure_acronym!="DGgr")[1:10,], subset(data,structure_acronym=="DGsg"|structure_acronym=="DGgr"))) + geom_bar(aes(x=structure_name, y=SOX11, fill=SOX11),stat='identity') + theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(hjust=0.5))+ labs(title="Regions with the top 10 expression:\nSOX11")

# DCX
data <-candidate.data[order(-candidate.data$DCX),]
## All regions
ggplot(data) + geom_bar(aes(x=structure_name, y=DCX, fill=DCX),stat='identity') + theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(hjust=0.5))
## Top 10 expression
ggplot(rbind(subset(data,structure_acronym!="DGsg" & structure_acronym!="DGgr")[1:10,], subset(data,structure_acronym=="DGsg"|structure_acronym=="DGgr"))) + geom_bar(aes(x=structure_name, y=DCX, fill=DCX),stat='identity') + theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(hjust=0.5))+ labs(title="Regions with the top 10 expression:\nDCX")

# MKI67
data <-candidate.data[order(-candidate.data$MKI67),]
## All regions
ggplot(data) + geom_bar(aes(x=structure_name, y=MKI67, fill=MKI67),stat='identity') + theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(hjust=0.5))
## Top 10 expression
ggplot(rbind(subset(data, structure_acronym!="DGsg" & structure_acronym!="DGgr")[1:10,], subset(data,structure_acronym=="DGsg"|structure_acronym=="DGgr"))) + geom_bar(aes(x=structure_name, y=MKI67, fill=MKI67),stat='identity') + theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1),plot.title = element_text(hjust=0.5)) + labs(title="Regions with the top 10 expression:\nMIK67")
```


# __[Approach 1: Unsupervised Clustering]__

## PCA: All Genes 
```{r,message=F,warning=F}
first.PC <-1; last.PC <-4

# Run PCA
  # full_PCA <- prcomp(exprs(ex_mac), scale.=T, center=F)
  # save(full_PCA, file="all.genes_PCA.rda")
  load("all.genes_PCA.rda")
# Get proportion of variance
per_PC1 <- round(summary(full_PCA)$importance[2,1]*100,2)
per_PC2 <- round(summary(full_PCA)$importance[2,2]*100,2)
# Create data.frame
full_PCA.data <- data.frame(full_PCA$rotation[,first.PC:last.PC])
full_PCA.data$Region <- pData(ex_mac)$structure_acronym
full_PCA.data$Age <- factor(pData(ex_mac)$age, c("E40","E50","E70","E80","E90","E120","0 mo","3 mo","12 mo","48 mo"),ordered=T)


# plot PCA
dat <-full_PCA.data
DGsg <-subset(dat,Region=="DGsg")
# By Region
### Plot
library(ggplot2)
ggplot(data=dat) + geom_point(aes(x=PC1, y=PC2, fill=Region, color=Region), size=2) +
  geom_point(data=DGsg, aes(x=PC1, y=PC2, colour=Region), size=4, shape=15) +
  geom_point(data=DGsg, aes(x=PC1, y=PC2), colour="black", size=4, shape=0) + labs(title="PCA: Colored by Region \n[SGZ samples outlined]", x=paste("PC1 (",per_PC1,"%)"), y=paste("PC2 (",per_PC2,"%)")) + theme(plot.title = element_text(hjust=0.5), legend.position="None")
## By Age
ggplot(data=dat) + geom_point(aes(x=PC1, y=PC2, fill=Age, color=Age), size=2) +
  geom_point(data=DGsg, aes(x=PC1, y=PC2,colour=Age), size=4, shape=15) +
  geom_point(data=DGsg, aes(x=PC1, y=PC2),colour="black", size=4, shape=0) + labs(title="PCA: Colored by Age \n[SGZ samples outlined]", x=paste("PC1 (",per_PC1,"%)"), y=paste("PC2 (",per_PC2,"%)")) + theme(plot.title = element_text(hjust=0.5))
# Get top PC1 loadings
PC1.load <-full_PCA$x[order(-abs(full_PCA$x[,1])),][,"PC1"]
write.csv(PC1.load[1:100],"PC1.allgenes.csv")
pander(PC1.load[1:10] , style='simple',justify='left')
```

## Unsupervised Clustering

__Interactive Heatmap Instructions:__ 
  + Hover cursor over cells to see info.
  + Click and drag a box over the area you want to zoom in on.
  + Click a column or row label to highlight just that column/row.
  + Double click to return to original view.
  
```{r,message=F,warning=F}
library(cluster); library(factoextra); library(magrittr)


# [1: Distance Methods]
unsup.clust_distance <- function(data){
res.dist <- get_dist(data, stand=T, method="pearson")
  library(d3heatmap) # use interactive heatmap instead
d3heatmap(res.dist, scale = "column", k_row=num_clusts, k_col=num_clusts)
#fviz_dist(res.dist, gradient=list(low="#00AFBB", mid="white", high="#FC4E07"))
}

# [2: Partitioning Methods]
# [[K-means clustering]]
unsup.clust_kmeans <-function(data.){
  #library("NbClust")
  #nb <- NbClust(data., distance="euclidean", min.nc=2, max.nc=10, method="kmeans")
  library("factoextra")
  # Conduct kmean using that number
  set.seed(123)
  km.res <- kmeans(data., num_clusts, nstart = 25) # Automatically extract # of clusters
  ## Visualize
  fviz_cluster(km.res, data = data.,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal(), repel=F)
}
  
# [[Partitioning Around Medoids (PAM) clustering]]
unsup.clust_PAM <-function(data.){
  library("cluster")
  pam.res <- pam(data., num_clusts)
  ## Visualize
  fviz_cluster(pam.res, data = data.,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal(), repel=T)
}


# [Hierarchical Methods]
# [[Hierarchical Clustering]]
unsup.clust_hierarchical <- function(data){
  library("factoextra")
  res.hc <- data %>%
    scale() %>%                    # Scale the data
    dist(method = "euclidean") %>% # Compute dissimilarity matrix
    hclust(method = "ward.D2")     # Compute hierachical clustering
  # Visualize using factoextra
  # Cut in 4 groups and color by groups
  fviz_dend(res.hc, k = num_clusts, # Cut in (#) groups
            cex = 0.5, repel=F, # label size
            palette="aaas",
            color_labels_by_k =T, # color labels by groups
            rect=T, # Add rectangle around groups
            type="circular", #"rectangle", "circular", "phylogenic".
            phylo_layout="layout.auto",# layout.auto, layout_with_drl, layout_as_tree, layout.gem, layout.mds, layout_with_lgl
            horiz=T, main="Unsupervised Hierarchical Clustering"
            )
  assign("hierarchical.output",res.hc)
}


# [[Hierarchical Clustering on Principal Components (HCPC)]]
unsup.clust_hcpc <- function(facto_data){
  library(FactoMineR)
  # Principal Component Analysis:
  res.pca <-  PCA(facto_data, graph=F)
    plot(res.pca)
  # Clustering, auto nb of clusters:
  hc <- HCPC(res.pca, nb.clust= -1)
    hc$call$t$nb.clust # Number of suggest clusters
  # Construct a hierarchical tree from a partition (with 10 clusters)
    # (useful when the number of individuals is very important)
    # hc2 <- HCPC(iris[,1:4], kk=10, nb.clust=-1)
}
```

### All Genes
```{r,message=F,warning=F}
# Setup data
    unsup.eset <-ex_mac[,pData(ex_mac)$age=="48 mo"]
    my_data <- t(exprs(unsup.eset))
    row.names(my_data) <-make.unique(as.character(pData(unsup.eset)$structure_name))
    # my_data <-full_PCA.data[,first.PC:last.PC]
    # row.names(my_data) <-make.unique(as.character(full_PCA.data$Region))
# Estimate  number of clusters
    # est.clusts <-fviz_nbclust(my_data, kmeans, method = "gap_stat") # [0]
      # save(est.clusts, file="estimated.clusters_all.genes.rda")
      load("estimated.clusters_all.genes.rda")
      est.clusts
      num_clusts=5 # Select number of clusters based on results
```
```{r,message=F,warning=F}
# Conduct unsupervised clustering
     unsup.clust_distance(my_data) # [1]
    # unsup.clust_kmeans(my_data) # [2]
    # unsup.clust_PAM(my_data) #[3]
    # unsup.clust_hierarchical(my_data) #[4]
    # unsup.clust_hcpc(my_data) # [5]
```

### Gene Subset: Mouse SGZ-enriched (Table S1)
```{r,message=F,warning=F}
# Setup data
unsup.eset <-ex_mac[row.names(exprs(ex_mac)) %in% s1.genes, pData(ex_mac)$age=="48 mo"] 
    my_data <- t(exprs(unsup.eset))
    row.names(my_data) <-make.unique(as.character(pData(unsup.eset)$structure_name))
    # my_data <-full_PCA.data[,first.PC:last.PC]
    # row.names(my_data) <-make.unique(as.character(full_PCA.data$Region))
# Estimate  number of clusters
    est.clusts_S1.genes <-fviz_nbclust(my_data, kmeans, method = "gap_stat") # [0]
    est.clusts_S1.genes
      num_clusts=3 # Select number of clusters based on results
```
```{r,message=F,warning=F}
# Conduct unsupervised clustering
    unsup.clust_distance(my_data) # [1]
    # unsup.clust_kmeans(my_data) # [2]
    # unsup.clust_PAM(my_data) #[3]
    # unsup.clust_hierarchical(my_data) #[4]
    # unsup.clust_hcpc(my_data) # [5]
```

### Gene Subset: 'Macaque Tan Module' genes only
```{r,message=F,warning=F}
# Setup data
unsup.eset <-ex_mac[row.names(exprs(ex_mac)) %in% Mac_mod.tan, pData(ex_mac)$age=="48 mo"]
    my_data <- t(exprs(unsup.eset))
    row.names(my_data) <-make.unique(as.character(pData(unsup.eset)$structure_name))
    # my_data <-full_PCA.data[,first.PC:last.PC]
    # row.names(my_data) <-make.unique(as.character(full_PCA.data$Region))
# Estimate  number of clusters
    est.clusts_S4.genes <-fviz_nbclust(my_data, kmeans, method = "gap_stat") # [0]
    est.clusts_S4.genes
      num_clusts=5 # Select number of clusters based on results
```
```{r,message=F,warning=F}
# Conduct unsupervised clustering
    unsup.clust_distance(my_data) # [1]
    # unsup.clust_kmeans(my_data) # [2]
    # unsup.clust_PAM(my_data) #[3]
    # unsup.clust_hierarchical(my_data) #[4]
    # unsup.clust_hcpc(my_data) # [5]
```


# __[Approach 2: Compare Expression Profiles]__
## Differential Expression

### All Genes
```{r,message=F,warning=F}
library(limma); library(statmod)
SGZ.samples <-exprs(ex_mac[,pData(ex_mac)$structure_acronym==c("DGsg")& pData(ex_mac)$age=="48 mo"])

region_list <- as.character(unique(pData(ex_mac)$structure_name))
DGE_results <- NULL
DGE_summary <- data.frame()
for(region in region_list){
   region.samples <-exprs( ex_mac[,pData(ex_mac)$structure_name==region & pData(ex_mac)$age=="48 mo"] )
   limma.data <-cbind(SGZ.samples, region.samples)
    fit <-lmFit(limma.data, design=
                 c(rep(1,ncol(SGZ.samples)), rep(-1,ncol(region.samples))) )
    fit <- eBayes(fit)
    tab <- topTable(fit, number=Inf)
   DGE_results[[paste("SGZ vs.",region)]] <- list(tab)
   # Create summary table
   sig <-dim(subset(tab, adj.P.Val<=0.05))[1]
   total <- dim(tab)[1]
   percent_DEGs <-round(dim(subset(tab, adj.P.Val<=0.05))[1] / 
     dim(tab)[1] *100,3)
   DGE_summary <- rbind(DGE_summary, 
                        data.frame(Test=paste("SGZ vs.", region), Sig_DEGs=sig,
                             Total_genes=total, Percent_DEGs=round(sig/total*100,3),
                             Total_logFC=sum(abs(tab$logFC)),
                             Total_t.stat=sum(abs(tab$t))) 
                        )
}

# Print written summary
no_DEGs <-length(subset(DGE_summary, Sig_DEGs==0)$Sig_DEGs)
total_regions <-length(DGE_summary$Test)
paste(no_DEGs,"/",total_regions,"adult brain regions contained no Differentially Expressed Genes compared to adult SGZ.")
# Results table
dge <-subset(DGE_summary,Test!="<NA>")
pander(dge[order(c(dge$percent_DEGs, dge$Total_logFC)),], style='simple', justify='left', split.table=Inf)
```

### Mouse SGZ-enriched genes (Table S1)
```{r,message=F,warning=F}
#Subset just the candidate genes and re-run t-tests
  ## Dividing.Cell genes (Miller et al. 2013, Table S1)
  gene.sub <- ex_mac[row.names(exprs(ex_mac)) %in% s1.genes, pData(ex_mac)$age=="48 mo"]
  
SGZ.samples <-exprs(gene.sub[,pData(gene.sub)$structure_acronym==c("DGsg")])

region_list <- as.character(unique(pData(gene.sub)$structure_name))
DGE_results <- NULL
DGE_summary <- data.frame()
for(region in region_list){
   region.samples <-exprs( gene.sub[,pData(gene.sub)$structure_name==region] )
   limma.data <-cbind(SGZ.samples, region.samples)
    fit <-lmFit(limma.data, design=
                 c(rep(1,ncol(SGZ.samples)), rep(-1,ncol(region.samples))) )
    fit <- eBayes(fit)
    tab <- topTable(fit, number=Inf)
   DGE_results[[paste("SGZ vs.",region)]] <- list(tab)
   # Create summary table
   sig <-dim(subset(tab, adj.P.Val<=0.05))[1]
   total <- dim(tab)[1]
   percent_DEGs <-round(dim(subset(tab, adj.P.Val<=0.05))[1] / 
     dim(tab)[1] *100,3)
   DGE_summary <- rbind(DGE_summary, 
                        data.frame(Test=paste("SGZ vs.", region), Sig_DEGs=sig,
                             Total_genes=total, Percent_DEGs=round(sig/total*100,3),
                             Total_logFC=sum(abs(tab$logFC)),
                             Total_t.stat=sum(abs(tab$t))) 
                        )
}

# Print written summary
no_DEGs <-length(subset(DGE_summary, Sig_DEGs==0)$Sig_DEGs)
total_regions <-length(DGE_summary$Test)
paste(no_DEGs,"/",total_regions,"adult brain regions contained no Differentially Expressed Genes compared to adult SGZ.")
# Results table
dge <-subset(DGE_summary,Test!="<NA>")
pander(dge[order(c(dge$percent_DEGs, dge$Total_logFC)),], style='simple', justify='left', split.table=Inf)
```

### Macaque Tan module genes (Table S4)
```{r,message=F,warning=F}
#Subset just the candidate genes and re-run t-tests
  ## Dividing.Cell genes (Miller et al. 2013, Table S4)
  gene.sub <- ex_mac[row.names(exprs(ex_mac)) %in% Mac_mod.tan, pData(ex_mac)$age=="48 mo"]
  
SGZ.samples <-exprs(gene.sub[,pData(gene.sub)$structure_acronym==c("DGsg")])

region_list <- as.character(unique(pData(gene.sub)$structure_name))
DGE_results <- NULL
DGE_summary <- data.frame()
for(region in region_list){
   region.samples <-exprs( gene.sub[,pData(gene.sub)$structure_name==region] )
   limma.data <-cbind(SGZ.samples, region.samples)
    fit <-lmFit(limma.data, design=
                 c(rep(1,ncol(SGZ.samples)), rep(-1,ncol(region.samples))) )
    fit <- eBayes(fit)
    tab <- topTable(fit, number=Inf)
   DGE_results[[paste("SGZ vs.",region)]] <- list(tab)
   # Create summary table
   sig <-dim(subset(tab, adj.P.Val<=0.05))[1]
   total <- dim(tab)[1]
   percent_DEGs <-round(dim(subset(tab, adj.P.Val<=0.05))[1] / 
     dim(tab)[1] *100,3)
   DGE_summary <- rbind(DGE_summary, 
                        data.frame(Test=paste("SGZ vs.", region), Sig_DEGs=sig,
                             Total_genes=total, Percent_DEGs=round(sig/total*100,3),
                             Total_logFC=sum(abs(tab$logFC)),
                             Total_t.stat=sum(abs(tab$t))) 
                        )
}

# Print written summary
no_DEGs <-length(subset(DGE_summary, Sig_DEGs==0)$Sig_DEGs)
total_regions <-length(DGE_summary$Test)
paste(no_DEGs,"/",total_regions,"adult brain regions contained no Differentially Expressed Genes compared to adult SGZ.")
# Results table
dge <-subset(DGE_summary,Test!="<NA>")
pander(dge[order(c(dge$percent_DEGs, dge$Total_logFC)),], style='simple', justify='left', split.table=Inf)
```




# __[Approach 3: Compare Network Modules]__

## WGCNA: Macaque Microarray (SGZ/GCL samples only, all ages)
Attempt to recreate networks from Miller et al. 2013
```{r,message=F,warning=F}
  library(WGCNA)
  library(flashClust)
allowWGCNAThreads() # Enable multi-threading

all.SGZ.GCL <- ex_mac[,pData(ex_mac)$structure_acronym %in% c("DGsg","DGgr")]
datExpr <- exprs(all.SGZ.GCL)
#net_corrs <- function(eset1, eset2){


## Soft threshold
powers = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14) 
sft = pickSoftThreshold(datExpr, networkType="signed", powerVector=powers) # t(net$MEs)
### Plot powers
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=0.9,col="red");
# this line corresponds to using an R^2 cut-off of h
  #abline(h=0.9,col="red")

## Run
POWER=14 # Same as Miller et al. 2013
# Assume defaults for everything not mentioned in Miller et al. 2013
   #net.mac.SGZ_GCL = blockwiseModules(t(datExpr), power=POWER, networkType="signed")
 #save(net.mac.SGZ_GCL, file="WGCNA_ABA.macaque_SGZ.GCL.rda")

#___________________________________________________________________________________#
# Plot WGCNA results
load("~/Desktop/Research/Signatures_of_Neurogensis/WGCNA_ABA.macaque_SGZ.GCL.rda")
# Plot
num_mods <- length(colnames(net.mac.SGZ_GCL$MEs))
moduleColorsAutomatic = WGCNA::labels2colors(net.mac.SGZ_GCL$colors)
mColors = moduleColorsAutomatic[net.mac.SGZ_GCL$blockGenes[[1]]]
plotDendroAndColors(net.mac.SGZ_GCL$dendrograms[[1]], colors=mColors, dendroLabels=F, 
                    groupLabels=c(paste("deepSplit = 2\n# modules =",num_mods)), addGuide=T,main="WGCNA Modules:\n Macaque SGZ & GCL samples")

# Get mutually exclusive modules assignments for each gene
mod_assignments <-data.frame(ModuleColors=net.mac.SGZ_GCL$colors, genes=row.names(datExpr))
mod_assignments <- mod_assignments[order(mod_assignments$ModuleColors),]
write.csv(mod_assignments, "module.gene.assignments_SGZ.GCL.csv")
```

### Rough '% overlap with genes in Tan module' metric
```{r,message=F,warning=F}
modules <-unique(mod_assignments$ModuleColors)
module.overlap_summary <-data.frame()

for(color in modules){
  one.module <-subset(mod_assignments,ModuleColors==color)
  percent_tan.overlap <-round(sum(Mac_mod.tan %in% as.character(one.module$genes)) / length(one.module$genes)*100,3)
  module.overlap_summary <- rbind(module.overlap_summary,data.frame(New_Module=color, New_Module_Size=length(one.module$genes), Percent_chance.overlap=length(one.module$genes)/dim(exprs(ex_mac))[1]*100, Percent_Tan.overlap=percent_tan.overlap))
}

pander::pander(module.overlap_summary[order(-module.overlap_summary$Percent_Tan.overlap),],style='simple',justify='left', split.table=Inf)
```



## WGCNA: Macaque Microarray (all regions, all genes, all ages)
Create new networks using all samples
```{r,message=F,warning=F}
 datExpr <- exprs(ex_mac)
#net_corrs <- function(eset1, eset2){
  library(WGCNA)
  library(flashClust)

## Soft threshold
powers = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14) 
sft = pickSoftThreshold(datExpr, networkType="signed", powerVector=powers) # t(net$MEs)
### Plot powers
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=0.9,col="red");
# this line corresponds to using an R^2 cut-off of h
  #abline(h=0.9,col="red")

## Run
POWER=14 # Same as Miller et al. 2013
# Assume defaults for everything not mentioned in Miller et al. 2013
  #net.mac = blockwiseModules(t(datExpr), power=POWER, networkType="signed", deepSplit=2)
# save(net.mac, file="WGCNA_ABA.macaque_all.samples.rda")

#___________________________________________________________________________________#
# Plot WGCNA results
load("~/Desktop/Research/Signatures_of_Neurogensis/WGCNA_ABA.macaque_all.samples.rda")
# Plot
num_mods <- length(colnames(net.mac$MEs))
moduleColorsAutomatic = WGCNA::labels2colors(net.mac$colors)
mColors = moduleColorsAutomatic[net.mac$blockGenes[[1]]]
plotDendroAndColors(net.mac$dendrograms[[1]], colors=mColors, dendroLabels=F, 
                    groupLabels=c(paste("deepSplit = 2\n# modules =",num_mods)), addGuide=T,main="Cluster Dendrogram")

# Get mutually exclusive modules assignments for each gene
mod_assignments <-data.frame(ModuleColors=net.mac$colors, genes=row.names(datExpr))
mod_assignments <-mod_assignments[order(mod_assignments$ModuleColors),]
write.csv(mod_assignments, "module.gene.assignments_all.regions.csv")
```

### Rough '% overlap with genes in Tan module' metric
```{r,message=F,warning=F}
modules <-unique(mod_assignments$ModuleColors)
module.overlap_summary <-data.frame()

for(color in modules){
  one.module <-subset(mod_assignments,ModuleColors==color)
  percent_tan.overlap <-round(sum(Mac_mod.tan %in% as.character(one.module$genes)) / length(one.module$genes)*100,3)
  module.overlap_summary <- rbind(module.overlap_summary,data.frame(New_Module=color, New_Module_Size=length(one.module$genes), Percent_chance.overlap=length(one.module$genes)/dim(exprs(ex_mac))[1]*100, Percent_Tan.overlap=percent_tan.overlap))
}

pander::pander(module.overlap_summary[order(-module.overlap_summary$Percent_Tan.overlap),],style='simple',justify='left', split.table=Inf)
```

* NOTES: There's almost no overlap between the new modules (derived from all macaque samples) and the tan module (derived from only macaque SGZ/GCL samples).


### Network Preservation Analysis
```{r,message=F,warning=F, eval=F}
# Subset all samples from SGZ or GCL
all.SGZ.GCL <- ex_mac[,pData(ex_mac)$structure_acronym %in% c("DGsg","DGgr")]
# Setup up inputs
multiExpr  = list(A1=list(data=t(exprs(ex_mac))), A2=list(data=t(exprs(all.SGZ.GCL))) )
multiColor = list(A1 = net.mac$colors)
# Run network preservation
mp <- modulePreservation(multiExpr, multiColor, referenceNetworks=1, verbose=3, networkType="signed", nPermutations=30, maxGoldModuleSize=100)
save(mp, file="network.preservation_all.samples.vs.GCL.SGZ.rda")
```

### Plot network preservation scores
```{r,message=F,warning=F}
load("~/Desktop/Research/Signatures_of_Neurogensis/network.preservation_all.samples.vs.GCL.SGZ.rda")
# Summary stats
stats <- mp$preservation$Z$ref.A1$inColumnsAlsoPresentIn.A2 # All stats output
pander::pander(stats[order(-stats[,2]),c(1:2)][1:10,],style='simple',justify='left')
 # Plot summary stats (Zsummary.pres)
  library(ggplot2)
ggplot(stats, aes(y=Zsummary.pres, x=factor(rownames(stats)),
                                 fill=factor(rownames(stats))) ) +
  geom_bar(stat="identity") + labs(title=paste("Network preservation scores"), y="Zsummary.pres", x="Network module") + theme(legend.position="none",axis.text.x = element_text(angle=45, hjust=1)) + scale_fill_manual(values=rownames(stats))
```




# __[Approach 4: Deconvolution]__

## Format scRNAseq reference database
* Redo with a more relevant sc-RNA-seq reference database
```{r,warning=F,message=F, eval=F}
library(EWCE)

###############################################################
## [1]: Default mouse hippocampus scRNAseq reference dataset ##
###############################################################
download.file("http://linnarssonlab.org/blobs/cortex/expression_mRNA_17-Aug-2014.txt",
    destfile="expression_mRNA_17-Aug-2014.txt") 
path = "expression_mRNA_17-Aug-2014.txt"
celltype_data = read_celltype_data(path)
load(file="~/Desktop/Research/Deconvolution_of_Evolution/celltype_data.rda")


###############################################################
## [2]: Shin et. al. 2015 scRNAseq reference dataset         ##
###############################################################
# First, download 'Series Matrix File(s)' from the GEO page (GSE71485) using the ARCHS4 Chrome plugin (to get the gene counts).
library(GEOquery)
gse <- getGEO(filename = "~/Desktop/GSE71485_GPL17021_series_matrix.txt.gz", destdir = ".", getGPL = F ) 
Shin <- read_celltype_data(path="~/Desktop/GSE71485_GPL17021_series_matrix.txt.gz")
celltype_data <- exprs(gse)
```

## 'Mouse SGZ-enriched' genes (Table S1)
### Setup data
```{r,message=F,warning=F}
# Set up gene list
example_genelist <- S1.genes
# Find mouse homologues of human genes
  data("mouse_to_human_homologs")
  m2h = unique(mouse_to_human_homologs[,c("HGNC.symbol","MGI.symbol")])
  mouse.hits = unique(m2h[m2h$MGI.symbol %in% example_genelist,"MGI.symbol"])
  mouse.bg  = unique(setdiff(m2h$MGI.symbol,mouse.hits))
  
paste(length(mouse.hits),"/",length(example_genelist),"candidate genes found in sc-RNA-seq reference database.")
```

### Sub-cell type level
```{r,message=F,warning=F}
# Bootstrap significance testing:
## 1) Without controlling for transcript length and GC content
full_results.sub = bootstrap.enrichment.test(sct_data=celltype_data,
                                         mouse.hits=mouse.hits,mouse.bg=mouse.bg,
                                         reps=10000, sub=T)
full_results.sub$results[order(full_results.sub$results$p),3:5][1:6,]
# Bootstrap plot
ewce.plot(full_results.sub$results,mtc_method="BH")


## 2) Controlling for transcript length and GC content (NOT WORKING)
### Need to convert gene list to human gene symbols
human.hits = unique(m2h[m2h$HGNC.symbol %in% example_genelist,"HGNC.symbol"])
human.bg = unique(setdiff(m2h$HGNC.symbol,human.hits))
# Bootstrap significance testing controlling for transcript length and GC content
#cont_results.sub = bootstrap.enrichment.test(sct_data=celltype_data,
 #                               human.hits=human.hits, human.bg=human.bg,
 #                                reps=10000, sub=T, geneSizeControl=T)
#ewce.plot(cont_results.sub$results,mtc_method="BH")
```

+ NOTES: EWCE deconvolution pretty nicely reconstructs cell types from 'mouse SGZ-enriched' genes.

### Cell type level
```{r,message=F,warning=F}
# 1) Bootstrap significance testing without controlling for transcript length and GC content
full_results = bootstrap.enrichment.test(sct_data=celltype_data,
                                         mouse.hits=mouse.hits,mouse.bg=mouse.bg,
                                         reps=10000, sub=F)
ewce.plot(full_results$results ,mtc_method="BH")

# 1) Controlling for transcript length and GC content (NOT WORKING)
#cont_results = bootstrap.enrichment.test(sct_data=celltype_data,human.hits=human.hits,human.bg=human.bg,reps=reps,sub=F,geneSizeControl=T)
#ewce.plot(cont_results$results,mtc_method="BH")
```

## 'Macaque tan module' genes (Table S4)
### Setup data
```{r,message=F,warning=F}
# Set up gene list
example_genelist <- Mac_mod.tan

library("EWCE")
load(file="~/Desktop/Research/Deconvolution_of_Evolution/celltype_data.rda")
# Find mouse homologues of human genes
  data("mouse_to_human_homologs")
  m2h = unique(mouse_to_human_homologs[,c("HGNC.symbol","MGI.symbol")])
  mouse.hits = unique(m2h[m2h$HGNC.symbol %in% example_genelist,"MGI.symbol"])
  mouse.bg  = unique(setdiff(m2h$MGI.symbol,mouse.hits))
  
paste(length(mouse.hits),"/",length(example_genelist),"candidate genes found in sc-RNA-seq reference database.")
```

### Sub-cell type level
```{r,message=F,warning=F}
# Bootstrap significance testing:
## 1) Without controlling for transcript length and GC content
full_results.sub = bootstrap.enrichment.test(sct_data=celltype_data,
                                         mouse.hits=mouse.hits,mouse.bg=mouse.bg,
                                         reps=10000, sub=T)
full_results.sub$results[order(full_results.sub$results$p),3:5][1:6,]
# Bootstrap plot
ewce.plot(full_results.sub$results,mtc_method="BH")


## 2) Controlling for transcript length and GC content (NOT WORKING)
### Need to convert gene list to human gene symbols
human.hits = unique(m2h[m2h$HGNC.symbol %in% example_genelist,"HGNC.symbol"])
human.bg = unique(setdiff(m2h$HGNC.symbol,human.hits))
# Bootstrap significance testing controlling for transcript length and GC content
#cont_results.sub = bootstrap.enrichment.test(sct_data=celltype_data,
 #                               human.hits=human.hits, human.bg=human.bg,
 #                                reps=10000, sub=T, geneSizeControl=T)
#ewce.plot(cont_results.sub$results,mtc_method="BH")
```

+ NOTES: EWCE deconvolution pretty nicely reconstructs cell types from 'mouse SGZ-enriched' genes.

### Cell type level
```{r,message=F,warning=F}
# 1) Bootstrap significance testing without controlling for transcript length and GC content
full_results = bootstrap.enrichment.test(sct_data=celltype_data,
                                         mouse.hits=mouse.hits,mouse.bg=mouse.bg,
                                         reps=10000, sub=F)
ewce.plot(full_results$results ,mtc_method="BH")

# 1) Controlling for transcript length and GC content (NOT WORKING)
#cont_results = bootstrap.enrichment.test(sct_data=celltype_data,human.hits=human.hits,human.bg=human.bg,reps=reps,sub=F,geneSizeControl=T)
#ewce.plot(cont_results$results,mtc_method="BH")
```

## Iteratively run EWCE on each region's DGE signature

### Run DGE: Each region vs. all other regions
```{r,warning=F,message=F}
samples <-ex_mac[,pData(ex_mac)$age=="48 mo"]

region_list <- as.character(unique(pData(samples)$structure_acronym))
DGE_results <- NULL
DGE_summary <- data.frame()
for(region in region_list){
    other.samples <-exprs( samples[,pData(samples)$structure_acronym!=region] )
   region.samples <-exprs( samples[,pData(samples)$structure_acronym==region] )
  # Run DGE
   limma.data <-cbind(region.samples, other.samples)
    fit <-lmFit(limma.data, design=
                 c(rep(1,ncol(region.samples)), rep(-1,ncol(other.samples))) )
    fit <- eBayes(fit)
    tab <- topTable(fit, number=Inf)
    tab <- tab[order(-abs(tab$logFC)),]
  
   DGE_results[[paste(region,"vs. all other regions")]] <- list(tab)
   # Create summary table
   sig <-dim(subset(tab, adj.P.Val<=0.05))[1]
   total <- dim(tab)[1]
   percent_DEGs <-round(sig / total *100,3)
   DGE_summary <- rbind(DGE_summary,data.frame(Test=paste(region,"vs. all other regions"),Sig_DEGs=sig, Total_genes=total, Percent_DEGs=round(sig/total*100,3),
                             Total_logFC=sum(abs(subset(tab, adj.P.Val<=0.05)$logFC)) ) 
                        )
}

# Print written summary
per_DEG <-length(subset(DGE_summary, percent_DEGs==100)$Sig_DEGs)
total_regions <-length(DGE_summary$Test)
paste(per_DEG,"/",total_regions,"adult brain regions have ALL genes differentially expressed.")
# Results table
pander::pander(subset(DGE_summary[order(c(DGE_summary$Percent_DEGs, -DGE_summary$Total_logFC)),], Test!="<NA>"),style='simple', justify='left', split.table=Inf)
```

### Run EWCE for each region
```{r,warning=F,message=F, echo=F}
# The first step is to load the data, obtain the MGI ids, sort the rows by t-statistic and then select the most up/down-regulated genes. 
  ## Example data
  #data(tt_alzh)
tt_results <- NULL
for(region in region_list){
# select Region
  #region <- "DGsg" 
  tt_data <-DGE_results[[paste(region,"vs. all other regions")]][[1]]
  tt_data$HGNC.symbol <-row.names(tt_data)
# Run analysis
tt_results[region] <-list(ewce_expression_data(sct_data=celltype_data, tt=tt_data))
# Plot
tt_results[[region]]$plot <-ewce.plot(tt_results[[region]]$joint_results)
tt_results[[region]]$plot
}

# Get specific plots
tt_results[["DGsg"]]$plot
tt_results[["DGgr"]]$plot
```

### Subset regions enriched in certain cell types
```{r,warning=F,message=F}
cell.type1 <-"astrocytes-ependymal"
cell.type2 <-"oligodendrocytes"
cell.type3 <-"interneurons"

enrichment.summary <-data.frame()
for(region in region_list){
  # 1st cell type
  if(subset(tt_results[[region]]$joint_results, CellType==cell.type1 &
            Direction=="Up")$p<=0.05){enriched1="Yes"} else{enriched1="No"}
  # 2nd cell type
    if(subset(tt_results[[region]]$joint_results, CellType==cell.type2 &
            Direction=="Up")$p<=0.05){enriched2="Yes"} else{enriched2="No"}
  # 3rd cell type
    if(subset(tt_results[[region]]$joint_results, CellType==cell.type3 &
            Direction=="Up")$p<=0.05){enriched3="Yes"} else{enriched3="No"}
  # Summarise
  enrichment.summary <-rbind(enrichment.summary, 
            data.frame(Region=region, Cell.Type=cell.type1, Enriched=enriched1), 
            data.frame(Region=region, Cell.Type=cell.type2, Enriched=enriched2),
            data.frame(Region=region, Cell.Type=cell.type3, Enriched=enriched3))
}

# Get only significantly enriched regions
enrich.summ <-subset(enrichment.summary, Enriched=="Yes")
pander(enrich.summ[order(enrich.summ$Region),], style='simple',justify='left')
```

# Jeremy WGCNA scripts
### WGCNA
```{r}
# Jeremy's Dropbox data
load("~/Desktop/Research/Signatures_of_Neurogensis/neurogenesisPaperData_ComBat_5_29_12.RData")
# ABA data
#load("~/Desktop/Research/Signatures_of_Neurogensis/ABA.macaque_preprocessed.v4.RData")

library(WGCNA); library(flashClust)
############################################################################# 
## Perform WGCNA on the macaque DG-gcl/DG-sgz data using ALL macaque genes ############################################################################### Choose the correct power for the network, and then find the TOM and clustering tree 
datExprMq2 =   2^t(datDG)#2^t(datExpr)#
pickSoftThreshold(datExprMq2,networkType="signed",powerVector= c(10,12,14,16,18))# Power SFT.R.sq slope 
powerMq2    = 14; AdjMatMq2 = adjacency(datExprMq2,power=powerMq2,type="signed"); diag(AdjMatMq2) = 0 
dissTOMq2   = 1-TOMsimilarity(AdjMatMq2, TOMType="signed");  collectGarbage();   geneTreeMq2 = flashClust(as.dist(dissTOMq2),method="average"); collectGarbage();   
## Plot the gene dendrograms and try out various parameters for tree cutting
mColorh=NULL; mName = NULL 
for (ds in c(0,1,2)) for (num in c(50,100,500))  {  
  tree = cutreeHybrid(dendro = geneTreeMq2, pamStage=FALSE, minClusterSize = num, deepSplit = ds, distM = dissTOMq2, cutHeight=0.99)  
  mColorh = cbind(mColorh,labels2colors(tree$labels));  
  mName   = cbind(mName, paste("ds =",ds,"- num =",num)) } 
pdf("deepSplit_tree_MacaqueAllGenes_cuts_final.pdf", height=10,width=20)
#x11()
plotDendroAndColors(geneTreeMq2, mColorh, mName, dendroLabels=FALSE, hang=0.03, addGuide=TRUE, guideHang=0.05, main="Gene dendrogram and module colors")
dev.off() 

# save(list=c("AdjMatMq2", "geneTreeMq2", "dissTOMq2", "geneTreeMq2"), file="Jeremy.data_SGZ.GCL_JeremyWGCNA.rda")
```

### Rough '% overlap with genes in Tan module' metric
```{r,message=F,warning=F}
# ABA data analyzed with Jeremy's long WGCNA method
load("~/Desktop/Research/Signatures_of_Neurogensis/WGCNA_tests/ABA_SGZ.GCL_JeremyWGCNA.rda")
# Jeremy's data analyzed with Jeremy's long WGCNA method

# Get mutually exclusive modules assignments for each gene
modules <- unique(mColorh[,1])
mod_assignments <-data.frame(ModuleColors=mColorh[,1], genes=row.names(datExpr))
mod_assignments <-mod_assignments[order(mod_assignments$ModuleColors),]
write.csv(mod_assignments, "module.gene.assignments_SGZ.GCL_JeremyWGCNA.csv")



modules <-unique(mod_assignments$ModuleColors)
module.overlap_summary <-data.frame()

for(color in modules){
  one.module <-subset(mod_assignments,ModuleColors==color)
  percent_tan.overlap <-round(sum(Mac_mod.tan %in% as.character(one.module$genes)) / length(one.module$genes)*100,3)
  module.overlap_summary <- rbind(module.overlap_summary,data.frame(New_Module=color, New_Module_Size=length(one.module$genes), Percent_chance.overlap=length(one.module$genes)/dim(exprs(ex_mac))[1]*100, Percent_Tan.overlap=percent_tan.overlap))
}

pander::pander(module.overlap_summary[order(-module.overlap_summary$Percent_Tan.overlap),],style='simple',justify='left', split.table=Inf)
```

```{r}
read_excel("~/Desktop/")
```


